---
title: " Worldwide People Life Dashboard" 
# author: "CID:02429075 "
# date: "5 May, 2024"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    vertical_layout: fill
    theme: yeti
runtime: shiny
---

``` {js}
// Inverse the color of navigation bar.
$('.navbar-inverse').removeClass('navbar-inverse').addClass('navbar-default');
```

```{r setup, include=FALSE}
# Load necessary packages. 
library(flexdashboard) # Dashboard package
library(highcharter) # Interactive data visualizations
library(viridis) # Color gradients
library(tidyverse) # Metapackge
library(countrycode) # Converting country names/codes
library(DT) # Displaying data tables
library(crosstalk) # Provides interactivity for HTML widgets
library(plotly) # Interactive data visualizations
library(dplyr)
library(tibble)
library(gridExtra)
library(shiny)
```

```{r}
# Define a new custom theme for Highcharts
new_custom_theme <- hc_theme(
  colors = c('#2b908f', '#f45b5b', '#90ee7e', '#7798BF', '#aaeeee', '#ff0066', '#eeaaee', '#55BF3B', '#DF5353', '#7798BF', '#aaeeee'),
  chart = list(
    backgroundColor = "#FAFAFA",
    plotBackgroundColor = "transparent",
    plotBorderColor = "#cccccc",
    plotBorderWidth = 1
  ),
  title = list(
    style = list(
      color = '#274e13',
      fontSize = "16px",
      fontFamily = "Verdana, sans-serif"
    )
  ),
  subtitle = list(
    style = list(
      color = '#4863A0',
      fontSize = "14px",
      fontFamily = "Verdana, sans-serif"
    )
  ),
  xAxis = list(
    gridLineColor = "#e6e6e6",
    gridLineWidth = 1,
    labels = list(
      style = list(
        color = "#333",
        fontSize = "12px"
      )
    ),
    lineColor = "#a6a6a6",
    tickColor = "#a6a6a6",
    title = list(
      style = list(
        color = "#123456",
        fontWeight = 'bold'
      )
    )
  ),
  yAxis = list(
    gridLineColor = "#e6e6e6",
    labels = list(
      style = list(
        color = "#333",
        fontSize = "12px"
      )
    ),
    lineColor = "#a6a6a6",
    minorGridLineColor = "#e6e6e6",
    tickWidth = 1,
    tickColor = "#a6a6a6",
    title = list(
      style = list(
        color = "#123456",
        fontWeight = 'bold'
      )
    )
  ),
  legend = list(
    backgroundColor = "rgba(0, 0, 0, 0.05)",
    itemStyle = list(
      color = '#333333',
      fontSize = "10px"
    ),
    itemHoverStyle = list(
      color = '#000'
    ),
    itemHiddenStyle = list(
      color = '#cccccc'
    )
  ),
  credits = list(
    enabled = FALSE
  ),
  tooltip = list(
    style = list(
      color = '#000000',
      fontWeight = 'bold'
    )
  )
)



```


```{r include=FALSE}

# Read in data. 
data1 <- read.csv('Timeuse.csv')

happiness_data <- read.csv('share-of-people-who-say-they-are-happy.csv') %>% 
  rename("Country" = "Entity",
         "Happiness" = "Happiness..Happy..aggregate.") %>% 
  select(-Code)
names(happiness_data) <- tolower(names(happiness_data))

annual_working_data <- read.csv('annual-working-hours-per-worker.csv') %>% 
  rename("Country" = "Entity",
         "ave_working_hour" = "Average.annual.working.hours.per.worker") %>%
  select(-Code)
names(annual_working_data) <- tolower(names(annual_working_data))

anxiety_data <- read.csv("share-who-report-lifetime-anxiety-or-depression.csv") %>% 
  rename("Country" = "Entity",
         "Anxiety" = "Share...Question..mh7a...Have.been.anxious.depressed...Answer..Yes...Gender..all...Age.group..all") %>% 
  select(-Code)
names(anxiety_data) <- tolower(names(anxiety_data))
anxiety_data[anxiety_data$country == "United States", "country"] = "United States of America"

anxiety_gap_data <- read.csv("anxiety-disorders-treatment-gap.csv") %>%
  select(-Code) %>%
  rename("Country" = "Entity",
         "potentially_adequate_treatment_conditional" = "Potentially.adequate.treatment..conditional",
         "other_treatments_conditional" = "Other.treatments..conditional",
         "untreated_conditional" = "Untreated..conditional") %>%
  select(-Year)
names(anxiety_gap_data) <- tolower(names(anxiety_gap_data))

anxiety_age_data <- read.csv("anxiety-disorders-prevalence-by-age.csv") %>%
  select(-Code) %>%
  select(-Anxiety.disorders..share.of.population....Sex..Both...Age..All.ages) %>%
  select(-Anxiety.disorders..share.of.population....Sex..Both...Age..Age.standardized) %>%
  
  rename("5-14_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..5.14.years",
         "15-19_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..15.19.years",
         "20-24_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..20.24.years",
         "25-29_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..25.29.years",
         "30-34_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..30.34.years",
         "35-39_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..35.39.years",
         "40-44_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..40.44.years",
         "45-49_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..45.49.years",
         "50-54_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..50.54.years",
         "55-59_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..55.59.years",
         "60-64_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..60.64.years",
         "65-69_age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..65.69.years",
         "70+age" = "Anxiety.disorders..share.of.population....Sex..Both...Age..70..years",
         "country" = "Entity",
         "year" = "Year")


life_exp_data <- read.csv("life-expectancy.csv") %>%
  select(-Code) %>%
  rename( "Country" = "Entity",
          "life_exp" = "Period.life.expectancy.at.birth...Sex..all...Age..0") 
names(life_exp_data) <- tolower(names(life_exp_data))
life_exp_data[life_exp_data$country == "United States", "country"] = "United States of America"

```




<style>
.colored {
  background-color: #FAFAFA;
}
</style>

Time Use 
========================================================================

Row {data-width=600 .colored }
-----------------------------------------------------------------------


### Worldwide Time use {.no-title}

```{r}

# Create a tibble grouped by country and category
country_category_time <- data1 %>%
  group_by(Country, Category) 
  

# Create an interactive bar plot
highchart() %>%
  hc_add_series(country_category_time, hcaes(x = Country, y = Time..minutes., group = Category), type = "bar") %>%
  hc_tooltip(borderWidth = 1.5, pointFormat = paste("Category: <b>{point.Category}</b> <br> Total Time: <b>{point.y}</b>")) %>%
  hc_title(text = "Total Time by Country and Category") %>%
  hc_subtitle(text = "Summary of activity time") %>%
  hc_xAxis(categories = unique(data1$Country),
           labels = list(step = 1),
           min = 0, max = 15,
           scrollbar = list(enabled = TRUE)) %>%
  hc_yAxis(title = list(text = "Total Time (minutes)"), min =0, max=1500) %>%
  hc_plotOptions(bar = list(stacking = "normal", 
                            pointPadding = 0.1, groupPadding = 0.1, borderWidth = 0.5)) %>%
  hc_credits(enabled = TRUE, text = "Data source: OECD Time use Database, Gender data Portal.", position = list(align = "center", verticalAlign = "bottom")) %>%
  hc_add_theme(new_custom_theme)



```


Columns {data-width=600 .colored }
-----------------------------------------------------------------------

### Trends in reported happiness by country {.no-title}
```{r}
selectInput('time_use_country', 
            'Country', 
            choices = unique(happiness_data$country), 
            selected = happiness_data$country[1])

# Create a tibble for our line plot. 

renderHighchart({
  happy_tibble <- reactive({
  subset(happiness_data, country == input$time_use_country)})

# Create line plot
highchart() %>% 
    hc_add_series(happy_tibble(), hcaes(x = year, y = happiness), type = "line", color = "#4863A0") %>%
    hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: <b>{point.x}</b> <br>", "Happiness: <b>{point.y}</b>")) %>%
    hc_title(text = paste("Trends in Reported Happiness in", input$time_use_country)) %>% 
    hc_subtitle(text = "Data source: Integrated Values Surveys (2022)") %>%
    hc_xAxis(title = list(text = "Year")) %>%
    hc_yAxis(title = list(text = "Average Happiness"),
             allowDecimals = FALSE) %>%
  hc_credits(enabled = TRUE, text = "Data source: Integrated Values Surveys (2022)    ", position = list(align = "center", verticalAlign = "top")) %>%
    hc_add_theme(new_custom_theme)
})



```


### Working hours {.no-title}
```{r}

selectInput('working_hours_country', 
            'Country', 
            choices = unique(annual_working_data$country), 
            selected = annual_working_data$country[1])

renderHighchart({
  working_tibble <- reactive({
  subset(annual_working_data, country == input$working_hours_country)})

# Create line plot
highchart() %>% 
    hc_add_series(working_tibble(), hcaes(x = year, y = ave_working_hour), type = "line", color = "#4863A0") %>%
    hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: <b>{point.x}</b> <br>","Country: <b>{point.country}</b><br>", "WorkingHours: <b>{point.y}</b>")) %>%
    hc_title(text = paste("Trends in working hours in", input$working_hours_country)) %>% 
    hc_subtitle(text = "Data source: Huberman & Minns (2007) and PWT 9.1 (2019)") %>%
    hc_xAxis(title = list(text = "Year")) %>%
    hc_yAxis(title = list(text = "Average working hours"),
             allowDecimals = FALSE) %>%
  hc_credits(enabled = TRUE, text = "Data source: Huberman & Minns (2007) and PWT 9.1 (2019)", position = list(align = "center", verticalAlign = "top")) %>%
    hc_add_theme(new_custom_theme)
})


```

Mental Health
=======================================================================


SideBar  {.sidebar}
-----------------------------------------------------------------------

```{r}
anx_age_tibble <- anxiety_age_data %>%
  pivot_longer(cols = c(`5-14_age`, `15-19_age`, `20-24_age`, `25-29_age`, `30-34_age`, `35-39_age`, `40-44_age`, `45-49_age`, `50-54_age`, `55-59_age`, `60-64_age`, `65-69_age`, `70+age`), names_to = "age", values_to = "value") %>%
  group_by(year, age)

selectInput('anxiety_country', 
            'Country for anxiety by age', 
            choices = unique(anx_age_tibble$country), 
            selected = anx_age_tibble$country[1])

############################################################
# Create a tibble for continent and sex.
an_gap_tibble <-  anxiety_gap_data %>%
  filter(country %in% c("High-income countries", "Upper-middle-income countries", "Lower-middle-income countries"))

# Reshape data to long format for pie chart
long_data <- an_gap_tibble %>%
  pivot_longer(cols = c(potentially_adequate_treatment_conditional, other_treatments_conditional, untreated_conditional), names_to = "treatment_res", values_to = "percentage")

#long_data <- subset(long_data, country == "High-income countries")

selectInput("income_level",
            "Type for the country:",
            choices = unique(long_data$country),
            selected =long_data$country[4])


```


Row {.tabset .tabset-fade data-width=500 .colored }
-----------------------------------------------------------------------

### Anxiety by age 
```{r}
# First, reshape the data from wide to long format
anx_age_tibble <- anxiety_age_data %>%
  pivot_longer(cols = c(`5-14_age`, `15-19_age`, `20-24_age`, `25-29_age`, `30-34_age`, `35-39_age`, `40-44_age`, `45-49_age`, `50-54_age`, `55-59_age`, `60-64_age`, `65-69_age`, `70+age`), names_to = "age", values_to = "value") %>%
  group_by(year, age)

#selectInput('anxiety_country', 
            #'Country', 
            #choices = unique(anx_age_tibble$country), 
            #selected = anx_age_tibble$country[1])


# Pick colors for the graph
age_color <- rev(plasma(13))

renderHighchart({
  anx_age_subtibble <- reactive({
  subset(anx_age_tibble, country == input$anxiety_country)})
  
  # Create the line plot
highchart() %>% 
  hc_add_series(anx_age_subtibble(), hcaes(x = year, y = value, group = age), type="line", color = age_color) %>%
  hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: <b>{point.x}</b> <br>","Age Group: <b>{point.age}</b><br>", "Persentage: <b>{point.y}%</b>")) %>%
  hc_title(text = "Percentage of anxiety by Age Group and Year") %>%
  hc_subtitle(text = paste("Trends in axiety in", input$anxiety_country)) %>%
  hc_xAxis(title = list(text = "Year")) %>%
  hc_yAxis(title = list(text = "Persentage"),
           allowDecimals = FALSE
           ) %>%
  hc_credits(enabled = TRUE, text = "Data source: IHME, Global Burden of Disease (2019)  ", position = list(align = "center", verticalAlign = "bottom")) %>%
  hc_add_theme(new_custom_theme)
})
  
```

### Treatment condition by country type

```{r}

category = c("High-income countries", "Upper-middle-income countries", "Lower-middle-income countries")

renderHighchart({
  long_data_tibble <- reactive({
    subset(long_data, country == input$income_level)
  })
  highchart() %>%
    hc_plotOptions(series = list(stacking = "percent")) %>%
    hc_add_series(data = long_data_tibble(), hcaes(x = treatment_res, y = percentage), type = "pie") %>%
    hc_xAxis(categories = category, label = list()) %>%
    hc_yAxis(title = list(text = "Percentage"), max = 100) %>%
    hc_legend(enabled = TRUE) %>%
    hc_tooltip(pointFormat =  paste( "Percentage: <b>{point.y}%</b>")) %>%
    hc_title(text = "Treatment and Untreated Rates by Country Income Level in 2017") %>%
    hc_subtitle(text = "Percentage Pie Chart") %>%
    hc_credits(enabled = TRUE, text = "Data source: World Mental Health surveys (2001–2015) ", position = list(align = "center", verticalAlign = "bottom")) %>%
    hc_add_theme(new_custom_theme)
  
})

```


Column {data-width=500}
-----------------------------------------------------------------------

### Worldwide anxiety {.no-title}
```{r}
highchart() %>% 
  hc_add_series_map(worldgeojson,
                       anxiety_data,
                       value = "anxiety",
                       joinBy = c("name", "country"),
                       name = "Anxiety") %>% 
  hc_tooltip(borderWidth = 1.5, headerFormat = "", valueSuffix = '% ') %>%
  hc_colorAxis(stops = color_stops()) %>% 
  hc_title(text = paste("Worldwide anxiety data in Year 2020")) %>%
  hc_credits(enabled = TRUE, text = "Data source: Wellcome Global Monitor (2021) ", position = list(align = "center", verticalAlign = "bottom")) %>%
  hc_add_theme(hc_theme_flat())

 
```
 


Life Expectancy
=======================================================================

SideBar  {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput('world_life_exp_year', 
            'Year', 
            choices = unique(life_exp_data$year), 
            selected = life_exp_data$year[1])
selectInput('world_life_exp_country', 
            'Country', 
            choices = unique(life_exp_data$country), 
            selected = life_exp_data$country[1])


```

Column {data-width=650}
-----------------------------------------------------------------------

### Worldwide Life Expectancy 
```{r}


renderHighchart({
    life_exp_tibble <- reactive({
    subset(life_exp_data, year == input$world_life_exp_year)
  })
    
    highchart() %>% 
  hc_add_series_map(worldgeojson,
                    life_exp_tibble(),
                    value = "life_exp",
                    joinBy = c("name", "country"),
                    name = "Life Expectancy") %>% 
  hc_tooltip(borderWidth = 1.5, valueSuffix = '') %>%
  hc_colorAxis(stops = color_stops()) %>% 
  hc_title(text = paste("Worldwide life expectancy data in Year ", input$world_life_exp_year)) %>%
  hc_credits(enabled = TRUE, text = "Data source: UN WPP (2022); HMD (2023); Zijdeman et al. (2015); Riley (2005)  ", position = list(align = "center", verticalAlign = "bottom")) %>%
  hc_add_theme(hc_theme_flat())

})

```


Column {data-width=350}
-----------------------------------------------------------------------

### Life Expectancy Trend
```{r}
renderHighchart({
    life_exp_tibble1 <- reactive({
    subset(life_exp_data, country == input$world_life_exp_country)
  })
    
# Create line plot.
highchart() %>% 
    hc_add_series(life_exp_tibble1(), hcaes(x = year, y = life_exp), type = "line") %>%
    hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, headerFormat = "", pointFormat = paste("Year: <b>{point.x}</b> <br> Life Expectancy: <b>{point.y}</b>")) %>%
    hc_title(text = paste("Life Expectancy by year in", input$world_life_exp_country)) %>% 
    hc_subtitle(text = "1950-2022") %>%
    hc_xAxis(title = list(text = "Year")) %>%
    hc_yAxis(title = paste("Life Expectancy in", input$world_life_exp_country),
             allowDecimals = FALSE) %>%
    hc_legend(enabled = FALSE) %>% 
  hc_credits(enabled = TRUE, text = "Data source: Same source with left plot", position = list(align = "center", verticalAlign = "bottom")) %>%
  hc_add_theme(hc_theme_flat())

})

```



About {data-icon="fa-info-circle"}
=======================================================================

Column {data-width=500}
-----------------------------------------------------------------------

### { .colored }

**Mental Health Hotlines:**

To find Mental Health hotlines in the UK, please view:

* https://www.nhs.uk/nhs-services/mental-health-services/

<br>





Column {data-width=500}
-----------------------------------------------------------------------

### { .colored }

**References:**  

The data set used to create the dashboard can be found at:   

* https://www.kaggle.com/russellyates88/suicide-rates-overview-1985-to-2016 

and was compiled from data from the following sources: 

* Wellcome Global Monitor (2021) – processed by Our World in Data. Retrieved from https://ourworldindata.org/mental-health

* Our World in Data based on Huberman & Minns (2007) and PWT 9.1 (2019) – processed by Our World in Data. Retrieved from https://ourworldindata.org/working-hours

* Data source: OECD Time use Database, Gender data Portal. Retrieved from https://docs.google.com/spreadsheets/d/1Cmav9S-bNUd-54HHXPRD1x-k3nDCceJPH_QXwQb1beU/edit#gid=2076567820

* Integrated Values Surveys (2022). Retrieved from https://ourworldindata.org/grapher/share-of-people-who-say-they-are-happy?tab=chart

* IHME, Global Burden of Disease (2020) – processed by Our World in Data. Retrieved from https://ourworldindata.org/grapher/anxiety-disorders-prevalence-by-age

* World Mental Health surveys (2001–2015). Retrieved from https://ourworldindata.org/grapher/anxiety-disorders-treatment-gap


